Bayesian Sparsification of Deep $\cplx$-valued networks

### 2020-02-07 paper/paper.tex
scattered typo fixes, and edits

% segue into bayesian inference
Bayesian inference is a very rich field of research with a large body of research and
techniques that enable practical quantification of predictive, parameter, or model
uncertainty and its propagation into solutions of the problems of interest.
%
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
which updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Bayesian methods with suitable modelling and prior assumptions, such
as Variational Dropout, can be used towards model sparsification
\citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015},
magnitude pruning, and $\ell_0$ regularization \citep{louizos_learning_2017} on large-scale
models. The latter is a probabilistic regularization method, which penalizes the expected
number of non-zero entries in $[0, 1]$-valued stochastic parameter masks with learnable
hard-concrete distribution with an atom at $0$ \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that although var-dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ on the studied architectures, pruning is preferable for
its simplicity and speed. They also observe that, despite dropout showing more variable
performance than magnitude pruning, it induces non-uniform sparsity throughout the model,
which has been shown by \citet{he_amc:_2018} to be essential for superior compression.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational sparsification and stochastic sparsification techniques we
extend Bayesian dropout for $\cplx$-valued networks. We assess the performance-compression
trade-off of the proposed Bayesian technique by conducting a large-scale numerical study
of $\cplx$-valued networks on two tasks: image recognition on MNIST-like and CIFAR10 datasets
and music transcription on MusicNet \citep{thickstun_learning_2017}. We reproduce the
state-of-the-art result by \citet{trabelsi_deep_2017} for $\cplx$-valued networks on the
latter with $\approx 5\%$ of remaining non-zero parameters.

@@

Being the ratio of the squared mean to the effective variance, $\alpha_j$ regulates the
relevance of the parameter $\omega_j$ it is associated to. Thus \citet{molchanov_variational_2017}
focus on the model sparsification capabilities of Variational Dropout, optimizing $\alpha$
for each individual parameter. Flexibility in specification of $q_\theta$ enables structured
dropout with shared $\alpha$ across groups of parameters $\omega$. Another variant of Bayesian
dropout, called \textit{Automatic Relevance Determination}, is developed in \citep{kharitonov_variational_2018}.
The key difference is that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper
Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes},
fits a prior distribution while performing variational inference. Maximizing \eqref{eq:elbo}
over $\tau$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

@@

Complex networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed a $\cplx$-differentiable real-valued function is
necessarily trivial by Cauchy-Riemann conditions. This issue is dealt with by employing Wirtinger,
or $\cplx\real$ calculus, which generalizes holomorphic calculus to non-holomorphic functions
of $\cplx$-argument \citep{adali_complex-valued_2011,boeddeker_computation_2019}. It satisfies
product and chain rules, respects complex conjugation and linearity, but most importantly the
$\cplx\real$-differential of a function on $\cplx$ coincides with its classical differential as
a function on $\real^2$. Essentially, it allows straightforward retrofiting of $\cplx$-valued
networks into existing $\real$ deep learning auto-differentiation frameworks. It was considered
as a basis for $\cplx$ version of backpropagation by \citet{benvenuto_complex_1992} and
\citet{trabelsi_deep_2017}, (see appendix~\ref{sub:wirtinger_calculus} for details).
% \todo{read on holomorphic nets}

@@

We consider two priors for Bayesian dropout: \verify{log-uniform} and $\cplx$-Gaussian empirical
Bayes \citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.
For a fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$,
the divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q \| \pi)
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = - \sum_{ij}
        \mathbb{H}(q(\omega_{ij}))
        + \mathbb{E}_{q(\omega_{ij})} \log{\pi(\omega_{ij})}
    \,.
\end{equation}
We omit subscripts ${ij}$ for brevity in the next sectons.

@@

on \eqref{eq:c-vd-kl-div}, it is possible to backpropagate through the divergence without

@@

Let's consider fully factorized proper $\cplx$-gaussian prior $
  \pi_\tau(\omega)
    \propto \mathcal{N}^{\cplx}\bigl(
      \omega \vert 0, \tau^{-1}, 0
    \bigr)
$. The per element divergence term in \eqref{eq:elbo-general-kl-div} is
\begin{equation}  \label{eq:emp-bayes-kl-div}
  KL(q \| \pi_\tau)
    = - 1 - \log{(\tau \sigma^2)}
      + \tau \bigl(
        \sigma^2 + \lvert \mu \rvert^2
      \bigr)
    \,.
\end{equation}
%
% This follows from the KL-divergence expression for two multivariate Gaussians
% $$
% % \mathbb{E}_{\omega\sim q_1(\omega)} \log \tfrac{q_1(\omega)}{q_2(\omega)}
% %   =
% KL(q_1\| q_2)
%   =
%   % - \tfrac12 \log\det{(2\pi e \Sigma_1)} + \tfrac12 \log\det{(2\pi \Sigma_2)}
%   - \tfrac12 \log \det{e I}
%   + \tfrac12 \log \frac{\det{\Sigma_2}}{\det{\Sigma_1}}
%   + \tfrac12 \tr{\bigl( \Sigma_2^{-1} \Sigma_1 \bigr)}
%   + \tfrac12 (\mu_1 - \mu_2)^\top \Sigma_2^{-1} (\mu_1 - \mu_2)
%   \,. $$
%
In empirical Bayes the prior adapts to the observed data, i.e. \eqref{eq:elbo} is optimized
w.r.t. $\tau$ of each weight's prior. Since divergence sits in \eqref{eq:elbo} with negative
sign, optimal $
  \tau^\ast = (\sigma^2 + \lvert \mu \rvert^2)^{-1}
$ is found by minimizing \eqref{eq:emp-bayes-kl-div}, which gives
\begin{equation}  \label{eq:emp-bayes-opt-kl}
  KL(q \| \pi_{\tau^\ast})
    % = - 1 - \log{({\tau^\ast} \sigma^2)}
    %   + {\tau^\ast} \sigma^2 + {\tau^\ast} \lvert \mu \rvert^2
    % = \log{((\sigma^2 + \lvert \mu \rvert^2) \tfrac1{\sigma^2})}
    = \log{\bigl(1 + \tfrac{\lvert \mu \rvert^2}{\sigma^2}\bigr)}
    = \log{\bigl(1 + \tfrac1\alpha \bigr)}
    \,.
\end{equation}

@@

% common experiment settings
Every experiment is replicated several times to take random effects into account, e.g.
random network initialization prior to training, randomly ordered batches during SGD.
Unless specified otherwise, every experiment uses global gradient clipping within an
$\ell_2$-norm ball of radius $0.5$, runs SGD with ADAM optimizer with the base learning
rate of $10^{-3}$. We note that since our experiments are on classification tasks,
$\cplx$-valued networks ultimately have to output a real-valued logit score. Following
\citet{wolter_complex_2018} and \citet{trabelsi_deep_2017}, the class scores are taken
as the real part of the $\cplx$ output.

@@

% on compresssion
Between \textit{``sparsify''} and \textit{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}). Since
the $q_\theta$ factorizes into univariate distributions, whether a $\cplx$ or $\real$ parameter
is zero or not is determined from $\alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}$: non-zero
if $\log \alpha \leq \tau$, zero otherwise. Figure~\ref{fig:hist__and__threshold__tradeoff}
(top) displays the general shape of the distribution of $\log \alpha$ on the Fashion-MNIST
dataset in sec.~\ref{sub:mnist_like_datasets}. This shape appears to be typical in every
experiment conducted in this study: there always is a single mode within the moderately
negative range, and one or two modes to the right of zero. In every experiment the trough
between the peaks of the negative mode and the first positive one is around zero. The
relative mass of each mode changes with $C$ and the higher the coefficient is, the more
mass is shifted to the right.
%
The bottom graph on figure~\ref{fig:hist__and__threshold__tradeoff} is constructed for
$C \in \{\tfrac1{20}, \frac1{200}\}$ before running the ``fine-tune'' stage. From the
relative positions of the compression curves on this plot and from the results of other
experiments we conclude that $C$ has much more substantial impact on the sparsity and
performance, than the choice of $\tau$. Furthermore, for $\tau > 0$ the performance quickly
saturates and while compression continues to drop.

@@

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. Since $W$ is $\real$- or a proper $\cplx$-Gaussian,
the random variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    % = \Pr\biggl(
    %   \frac{k \lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
    %     \leq \frac{\delta^2 k}{\alpha}
    % \biggr)
    = \chi^2_k\biggl(\frac{\delta^2 k}{\alpha}\biggr)
  \,.
\end{equation}
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. In our opinion, $\tau = 3$
\citep{molchanov_variational_2017,kingma_variational_2015} retains parameters with randomness
that significantly exceeds the magnitude of the mean, although $\mu$ is typically used as
a stand-in for $W$ when the network runs in inference mode. Given that $q_\theta$ is merely
an approximation, we pick $\tau = -\tfrac12$ to exclude noisy parameters and encourage
higher sparsity.

@@

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is computed based on the number of floating point values needed
to specify the model and equals $
  \frac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros after sparsification, $n_\mathtt{par}$
is the total number of floats. Thus each parameter is counted once in a $\real$-valued
network, but twice in a $\cplx$-network, being a pair of real-imaginary floats. Each model
has incompressible parameters, such as layer biases, shift and scaling from $\real$-and
$\cplx$-batch normalization layers, if present.

@@

% first experiment: 40-75-40 with fast scheduler
First we compare the $\real$-valued networks against their $\cplx$-valued counterparts with
approximately half as many parameters as suggested by \citet{monning_evaluation_2018}. In
this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base learning
rate is scaled after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.

% discuss the plots describing how the test-performance grows with compression and then drops if the model is too sparse
Plots on figure~\ref{fig:mnist-like__trade-off} show the results on the ``MNIST'' dataset for
$\tau\in\{-\tfrac12, \tfrac32\}$ and $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$ replicated $5$ times using the method from sec.~\ref{ssub:variational_dropout} (VD)
and~\ref{ssub:empirical_bayes} (ARD), respectively (results for other MNIST-like datasets
are in the appendix).

% which is better ARD or VD
It is worth pointing out that the prior used in ARD models offers slightly more favourable
compression-accuracy values, matching the established perception that log-uniform prior
is too harsh for sparsification tasks \citep{kharitonov_variational_2018}.

@@

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__musicnet__trade-off/musicnetram__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Average Percision frontier for the convolutional model of \citet{trabelsi_deep_2017}.
    Shaded horiontal region represents the $\mu\pm\sigma$ bands for the performance of the
    uncompressed model.
    \textit{top}: Original model ($k=6$) compressed with \textbf{\color{blue} VD}
    (\ref{ssub:variational_dropout}) and \textbf{\color{orange} ARD} (\ref{ssub:empirical_bayes});
    \textit{bottom}: The model of \citet{trabelsi_deep_2017} with kernel size $3$
    compressed with \textbf{\color{blue} VD} (\ref{ssub:variational_dropout}).
  }
  \label{fig:musicnet__trade-off}
\end{figure}

@@

If we assume a mean field $\cplx$-Gaussian approximate posteriror for $
  W \in \cplx^{n\times m}
$, then
\begin{equation}  \label{eq:c-gauss-vi-general-vec}
  \vec{W}
    \sim \mathcal{N}^{\cplx}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagonal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick-appendix}
  y_i
    \sim \mathcal{N}^{\cplx}
      \bigl(
        e_i^\top \mu x + b_i,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
Hence \eqref{eq:cplx-gauss-trick} folows.

@@

Natural identification of $\cplx$ with $\real\times \real$, storing the real and imaginary
parts in interleaved format, emulation of $\cplx$-algebra using $\real$-valued arithmetic,
and Wirtinger calculus make it possible to reuse $\real$-backpropagation and existing
auto-differentiation frameworks \citep{trabelsi_deep_2017}.

>>>>

% segue into bayesian inference
Bayesian inference is a very rich field of research with a large body of research and
techniques that enable practical quantification of predictive, parameter, or model
uncertainty and its propagation into solutions of the problems of interest.
%
Dropout resembles averaging models with shared parameters: each model is weighted equally
since masks are randomly sampled, unlike Bayesian model averaging, which weighs according to
the posterior distribution. \citet{kingma_variational_2015} specifically consider dropout
as a Bayesian Inference technique: a principled framework of reasoning about uncertainty,
which updates prior beliefs in accordance with the likelihood of empirical observations
into posterior belief. Bayesian methods with suitable modelling and prior assumptions, such
as Variational Dropout, can be used towards model sparsification
\citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.

% (sparsity tech large-scale models) gale-elsen-hooker.tex
\citet{gale_state_2019} compare variational dropout \citep{kingma_variational_2015},
magnitude pruning, and $\ell_0$ regularization \citep{louizos_learning_2017} on large-scale
models. The latter is a probabilistic regularization method, which penalizes the expected
number of non-zero entries in $[0, 1]$-valued stochastic parameter masks with learnable
hard-concrete distribution with an atom at $0$ \citep{maddison_concrete_2016,jang_categorical_2017}.
Their results suggest that although variational dropout may achieve good accuracy-sparsity balance
and outperform pruning and $\ell_0$ on the studied architectures, pruning is preferable for
its simplicity and speed. They also observe that, despite dropout showing more variable
performance than magnitude pruning, it induces non-uniform sparsity throughout the model,
which has been shown by \citet{he_amc:_2018} to be essential for superior compression.
% Magnitude pruning induces user-specified sparsity distributions.
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
%

% our contribution
Inspired by variational and stochastic sparsification techniques we extend Bayesian dropout
for $\cplx$-valued networks. We assess the performance-compression trade-off of the proposed
Bayesian technique by conducting a large-scale numerical study of $\cplx$-valued networks
on two tasks: image recognition on MNIST-like and CIFAR10 datasets and music transcription
on MusicNet \citep{thickstun_learning_2017}. We reproduce the state-of-the-art result by
\citet{trabelsi_deep_2017} for $\cplx$-valued networks on the latter with $\approx 5\%$
of remaining non-zero parameters.

@@

Being the ratio of the squared mean to the effective variance, $\alpha_j$ regulates the
relevance of the parameter $\omega_j$ it is associated to. Thus \citet{molchanov_variational_2017}
focus on the model sparsification capabilities of Variational Dropout, optimizing $\alpha$
for each individual parameter. Flexibility in specification of $q_\theta$ enables structured
dropout with shared $\alpha$ across groups of parameters $\omega$. Another Bayesian dropout
method, called \textit{Automatic Relevance Determination}, is developed in \citep{kharitonov_variational_2018}.
The key difference is that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper
Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes},
fits a prior distribution while performing variational inference. Maximizing \eqref{eq:elbo}
over $\tau$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta \| \pi)
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

@@

Complex networks are, in general, non-holomorphic, i.e. not $\cplx$-differentiable, which is
exacerbated by the loss being real-valued. Indeed a $\cplx$-differentiable real-valued function is
necessarily trivial by Cauchy-Riemann conditions. This issue is dealt with by employing Wirtinger,
or $\cplx\real$ calculus, which generalizes holomorphic calculus to non-holomorphic functions
of $\cplx$-argument \citep{adali_complex-valued_2011,boeddeker_computation_2019}. It satisfies
product and chain rules, respects complex conjugation and linearity, but most importantly the
$\cplx\real$-differential of a function on $\cplx$ coincides with its classical differential as
a function on $\real^2$. Essentially, it allows straightforward retrofiting of $\cplx$-valued
networks into existing $\real$ deep learning auto-differentiation frameworks. It was considered
as a basis for $\cplx$ version of back-propagation by \citet{benvenuto_complex_1992} and
\citet{trabelsi_deep_2017}, (see appendix~\ref{sub:wirtinger_calculus} for details).
% \todo{read on holomorphic nets}

@@

We consider two priors for Bayesian dropout: \verify{log-uniform} and $\cplx$-Gaussian empirical
Bayes \citep{kingma_variational_2015,molchanov_variational_2017,kharitonov_variational_2018}.
For a fully factorized approximation $q(\omega)$ and factorized prior belief $\pi(\omega)$,
the divergence term in ELBO \eqref{eq:elbo} is
\begin{equation}  \label{eq:elbo-general-kl-div}
  KL(q \| \pi)
    % = \mathbb{E}_{\oemga \sim q(\oemga)} \log \tfrac{q(\oemga)}{\pi(\oemga)}
    % = \sum_{ij} \mathbb{E}_{\omega_{ij} \sim q(\omega_{ij})}
    %   \log \tfrac{q(\omega_{ij})}{\pi(\omega_{ij})}
    % = \sum_{ij} KL\bigl( q(\omega_{ij}) \| \pi(\omega_{ij}) \bigr)
    = - \sum_{ij}
        \mathbb{H}(q(\omega_{ij}))
        + \mathbb{E}_{q(\omega_{ij})} \log{\pi(\omega_{ij})}
    \,.
\end{equation}
We omit subscripts ${ij}$ for brevity in the next sections.

@@

on \eqref{eq:c-vd-kl-div}, it is possible to back-propagate through the divergence without

@@

Let's consider fully factorized proper $\cplx$-Gaussian prior $
  \pi_\tau(\omega)
    \propto \mathcal{N}^{\cplx}\bigl(
      \omega \vert 0, \tau^{-1}, 0
    \bigr)
$. The per element divergence term in \eqref{eq:elbo-general-kl-div} is
\begin{equation}  \label{eq:emp-bayes-kl-div}
  KL(q \| \pi_\tau)
    = - 1 - \log{(\tau \sigma^2)}
      + \tau \bigl(
        \sigma^2 + \lvert \mu \rvert^2
      \bigr)
    \,.
\end{equation}
%
% This follows from the KL-divergence expression for two multivariate Gaussians
% $$
% % \mathbb{E}_{\omega\sim q_1(\omega)} \log \tfrac{q_1(\omega)}{q_2(\omega)}
% %   =
% KL(q_1\| q_2)
%   =
%   % - \tfrac12 \log\det{(2\pi e \Sigma_1)} + \tfrac12 \log\det{(2\pi \Sigma_2)}
%   - \tfrac12 \log \det{e I}
%   + \tfrac12 \log \frac{\det{\Sigma_2}}{\det{\Sigma_1}}
%   + \tfrac12 \tr{\bigl( \Sigma_2^{-1} \Sigma_1 \bigr)}
%   + \tfrac12 (\mu_1 - \mu_2)^\top \Sigma_2^{-1} (\mu_1 - \mu_2)
%   \,. $$
%
In Empirical Bayes the prior adapts to the observed data, i.e. \eqref{eq:elbo} is optimized
w.r.t. $\tau$ of each weight's prior. Since divergence sits in \eqref{eq:elbo} with negative
sign, optimal $
  \tau^\ast = (\sigma^2 + \lvert \mu \rvert^2)^{-1}
$ is found by minimizing \eqref{eq:emp-bayes-kl-div}, which gives
\begin{equation}  \label{eq:emp-bayes-opt-kl}
  KL(q \| \pi_{\tau^\ast})
    % = - 1 - \log{({\tau^\ast} \sigma^2)}
    %   + {\tau^\ast} \sigma^2 + {\tau^\ast} \lvert \mu \rvert^2
    % = \log{((\sigma^2 + \lvert \mu \rvert^2) \tfrac1{\sigma^2})}
    = \log{\bigl(1 + \tfrac{\lvert \mu \rvert^2}{\sigma^2}\bigr)}
    = \log{\bigl(1 + \tfrac1\alpha \bigr)}
    \,.
\end{equation}

@@

% common experiment settings
Every experiment is replicated several times to take random effects into account, e.g.
random network initialization prior to training, randomly ordered batches during SGD.
Unless specified otherwise, every experiment uses global gradient clipping within an
$\ell_2$-norm ball of radius $0.5$, runs SGD with ADAM optimizer with the base learning
rate of $10^{-3}$. We note that since our experiments are on classification tasks,
$\cplx$-valued networks ultimately have to output a real-valued logit score. Following
\citet{wolter_complex_2018} and \citet{trabelsi_deep_2017}, the class scores are taken
as real parts of the $\cplx$ output.

@@

% on compresssion
Between \textit{``sparsify''} and \textit{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}). Since
the $q_\theta$ factorizes into univariate distributions, a $\cplx$ or $\real$ parameter is
considered non-zero if $\log \alpha \leq \tau$ for $
  \alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}
$, or zero otherwise. Figure~\ref{fig:hist__and__threshold__tradeoff} (top) displays
the general shape of the distribution of $\log \alpha$ on the Fashion-MNIST dataset in
sec.~\ref{sub:mnist_like_datasets}. There always is a single mode within the moderately
negative range, and one or two modes to the right of zero, and the trough between the
peaks of the negative mode and the first positive one is around zero. The relative mass
of each mode changes with $C$ and the higher the coefficient is, the more mass is shifted
to the right.
%
The bottom graph on figure~\ref{fig:hist__and__threshold__tradeoff} is constructed
for $C \in \{\tfrac1{20}, \frac1{200}\}$ before running the ``fine-tune'' stage. From
\eqref{eq:elbo_with_coef}, the relative positions of the compression curves on this
plot and the results of other experiments, it can be concluded that $C$ has much more
substantial impact on the sparsity and performance, than the choice of $\tau$. Furthermore,
for $\tau > 0$ the performance quickly saturates and while compression continues to
drop.

@@

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining \textit{``non-zero''} parameters
are within $\delta$ relative tolerance of their mean with high probability under the
approximate posterior $q_\theta(W)$. This ensures that stochastic parameters are sufficiently
concentrated around $\mu$, which is used as a stand-in value for $W$, when the network
is used in inference mode. Since $W$ is $\real$- or a proper $\cplx$-Gaussian, the random
variable $
  \tfrac{k \lvert W - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$), whence
\begin{equation} \label{eq:prob_relevant}
  \Pr\bigl(
    \lvert w - \mu \rvert \leq \delta \lvert \mu \rvert
  \bigr)
    % = \Pr\biggl(
    %   \frac{k \lvert w - \mu \rvert^2}{\alpha \lvert \mu \rvert^2}
    %     \leq \frac{\delta^2 k}{\alpha}
    % \biggr)
    = \chi^2_k\biggl(\frac{\delta^2 k}{\alpha}\biggr)
  \,.
\end{equation}
For a generous $\delta = 50\%$ tolerance all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. We pick $\tau = -\tfrac12$
to exclude noisy parameters and encourage higher sparsity, at the same time staying aware
that $q_\theta$ is merely an approximation. In comparison, $\tau = 3$ is commonly used as
the threshold \citep{molchanov_variational_2017,kingma_variational_2015}.

@@

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is computed based on the number of floating point values needed
to specify the model and equals $
  \frac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros after sparsification, $n_\mathtt{par}$
is the total number of floats. Thus each parameter is counted once in a $\real$-valued
network, but twice in a $\cplx$-network, being a pair of real-imaginary floats. Each model
has incompressible parameters, such as layers' biases, shift and scaling from $\real$-and
$\cplx$-batch normalization layers, if present.

@@

% first experiment: 40-75-40 with fast scheduler
First we compare the $\real$-valued networks against their $\cplx$-valued counterparts with
approximately half as many parameters as suggested by \citet{monning_evaluation_2018}. In
this experiment stages last for $40$, $75$ and $40$ epochs, respectively, and the base learning
rate is scaled after $5$, $10$, and $20$-th epoch by $\tfrac1{10}$, $\tfrac1{20}$ and $\tfrac1{100}$,
respectively.
%
Plots~\ref{fig:mnist-like__trade-off} depict performance-compression curve on MNIST-like
datasets for the studied models and methods from sec.~\ref{ssub:variational_dropout} (VD)
and~\ref{ssub:empirical_bayes} (ARD). Besides that we tried \texttt{raw} or \texttt{fft}
input features, varied $
  C \in \{\tfrac3{2^{k+1}}\colon k=1..16\}
$, and repeated the experiment $4$ times.

% observation: raw features -- alright

fft features with wide dense network exhibit improving performance with the compression rate
up to a certain limit.

% discuss the plots describing how the test-performance grows with compression and then drops if the model is too sparse
% which is better ARD or VD
It is worth pointing out that the prior used in ARD models offers slightly more favourable
compression-accuracy values, matching the established perception that log-uniform prior
is too harsh for sparsification tasks \citep{kharitonov_variational_2018}.

@@

\begin{figure}[!ht]
  \centering
  \begin{subfigure}[b]{1.\textwidth}  % imcl2019-style swears at this
    \centering
    \includegraphics[width=\linewidth]{../assets/figure__musicnet__trade-off/musicnetram__fft__-0.5.pdf}
  \end{subfigure}
  \caption{%
    Performance-compression frontier for the deep network of \citet{trabelsi_deep_2017}.
    Shaded horizontal region represents the $\mu\pm\sigma$ bands for the performance of the
    uncompressed model. \textbf{\color{blue} VD} and \textbf{\color{orange} ARD} denote
    methods \ref{ssub:variational_dropout}) ans \ref{ssub:empirical_bayes}, respectively.
    The \textbf{\color{pink} version} of the network with kernel $3$ in every convolution compressed with VD
    is (\ref{ssub:variational_dropout}).
  }
  \label{fig:musicnet__trade-off}
\end{figure}

@@

If we assume a mean field $\cplx$-Gaussian approximate posterior for $
  W \in \cplx^{n\times m}
$, then
\begin{equation}  \label{eq:c-gauss-vi-general-vec}
  \vec{W}
    \sim \mathcal{N}^{\cplx}_{nm} \bigl(
      \vec{\mu}, \diag{\vec{\Sigma}}, \diag{\vec{C}}
    \bigr)
  \,,
\end{equation}
where mean, \textit{variance} and relation are given by matrices $
  \mu, \Sigma, C \in \cplx^{n\times m}
$ with $\Sigma_{ij} \geq 0$, and $
  \lvert C_{ij} \rvert^2 \leq \Sigma_{ij}
$. For any $x \in \cplx^m$ and $b\in \cplx^n$ $
  y = W x + b
    = (I_n \otimes x^\top) \vec{W} + b
$, whence the covariance and relation matrices of $y$ are
\begin{align}
  (I \otimes x^\top) \diag{\vec{\Sigma}} (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (I \otimes x^\top) e_i \otimes e_j
    %   \Sigma_{ij} (e_i \otimes e_j)^\top (I \otimes x^\top)^{\hop}
    % = \sum_{ij} (e_i \otimes x^\top e_j)
    %   \Sigma_{ij} (e_i \otimes x^\top e_j)^{\hop}
    % = \sum_{ij} (e_i \otimes x_j) \Sigma_{ij} (e_i^\top \otimes \conj{x}_j)
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m \Sigma_{ij} x_j \conj{x}_j
    \,,  \label{eq:cn-gauss-lrt-cov} \\
  (I \otimes x^\top) \diag{\vec{C}} (I \otimes x^\top)^\top
    % = \sum_{ij} (e_i \otimes x_j) C_{ij} (e_i \otimes x_j)^\top
    & = \sum_{i=1}^n (e_i e_i^\top) \sum_{j=1}^m C_{ij} x_j^2
    \,.  \label{eq:cn-gauss-lrt-rel}
\end{align}
Since \eqref{eq:cn-gauss-lrt-cov} and \eqref{eq:cn-gauss-lrt-rel} are diagonal, $(y_i)_{i=1}^n$
are independent whence \eqref{eq:cn-affine} implies
\begin{equation}  \label{eq:cplx-gauss-trick-appendix}
  y_i
    \sim \mathcal{N}^{\cplx}
      \bigl(
        e_i^\top \mu x + b_i,
        \, \sum_{j=1}^m \Sigma_{ij} \lvert x_j \rvert^2,
        \, \sum_{j=1}^m C_{ij} x_j^2
      \bigr)
    \,.
\end{equation}
Hence \eqref{eq:cplx-gauss-trick} follows.

@@

Natural identification of $\cplx$ with $\real\times \real$, storing the real and imaginary
parts in interleaved format, emulation of $\cplx$-algebra using $\real$-valued arithmetic,
and Wirtinger calculus make it possible to reuse $\real$ back-propagation and existing
auto-differentiation frameworks \citep{trabelsi_deep_2017}.
